### 2016-05-11 poster/poster.tex
Start on a poster

% vim: set spell spelllang=en tw=100 et sw=4 sts=4 :

\documentclass[a0paper]{tikzposter}

\usepackage{complexity}
\usepackage{wrapfig}
\usepackage{microtype}
\usepackage{gnuplot-lua-tikz}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{sfmath}

\usepackage{lmodern}
\renewcommand*\familydefault{\sfdefault}
\usepackage[T1]{fontenc}

\title{Heuristics and Really Hard Instances for Subgraph Isomorphism Problems}
\author{Ciaran McCreesh, Patrick Prosser and James Trimble}
\institute{University of Glasgow, Glasgow, Scotland}
\titlegraphic{\includegraphics[keepaspectratio=true,scale=3.5]{UoG_keyline.pdf}}

\settitle{
    \begin{tikzpicture}
        \node (T) [inner sep=0pt] {\begin{minipage}{\linewidth}
                \color{titlefgcolor}
                {\bfseries \Huge \hspace*{10mm}Heuristics and Really Hard Instances for \\
                \hspace*{10mm}Subgraph Isomorphism Problems \par}
                \vspace*{1em}
                {\Large {\bfseries \hspace{10mm}\@author}, \@institute}
        \end{minipage}};

        \node at (T.east) [anchor=center, inner sep=0pt, xshift=-12cm] {\@titlegraphic};
    \end{tikzpicture}
}

% University of Glasgow standard colours
\definecolor{uofguniversityblue}{rgb}{0, 0.219608, 0.396078}

\definecolor{uofgheather}{rgb}{0.356863, 0.32549, 0.490196}
\definecolor{uofgaquamarine}{rgb}{0.603922, 0.72549, 0.678431}
\definecolor{uofgslate}{rgb}{0.309804, 0.34902, 0.380392}
\definecolor{uofgrose}{rgb}{0.823529, 0.470588, 0.709804}
\definecolor{uofgmocha}{rgb}{0.709804, 0.564706, 0.47451}

\definecolor{uofglawn}{rgb}{0.517647, 0.741176, 0}
\definecolor{uofgcobalt}{rgb}{0, 0.615686, 0.92549}
\definecolor{uofgturquoise}{rgb}{0, 0.709804, 0.819608}
\definecolor{uofgsunshine}{rgb}{1.0, 0.862745, 0.211765}
\definecolor{uofgpumpkin}{rgb}{1.0, 0.72549, 0.282353}
\definecolor{uofgthistle}{rgb}{0.584314, 0.070588, 0.447059}
\definecolor{uofgpillarbox}{rgb}{0.701961, 0.047059, 0}
\definecolor{uofglavendar}{rgb}{0.356863, 0.301961, 0.580392}

\definecolor{uofgsandstone}{rgb}{0.321569, 0.278431, 0.231373}
\definecolor{uofgforest}{rgb}{0, 0.317647, 0.2}
\definecolor{uofgburgundy}{rgb}{0.490196, 0.133333, 0.223529}
\definecolor{uofgrust}{rgb}{0.603922, 0.227451, 0.023529}

\definecolorstyle{UofG}{
}{
    % Background Colors
    \colorlet{backgroundcolor}{uofgsandstone!80!white}
    \colorlet{framecolor}{black}
    % Title Colors
    \colorlet{titlefgcolor}{white}
    \colorlet{titlebgcolor}{uofguniversityblue}
    % Block Colors
    \colorlet{blocktitlebgcolor}{white}
    \colorlet{blocktitlefgcolor}{uofguniversityblue}
    \colorlet{blockbodybgcolor}{white}
    \colorlet{blockbodyfgcolor}{black}
    % Innerblock Colors
    \colorlet{innerblocktitlebgcolor}{uofguniversityblue}
    \colorlet{innerblocktitlefgcolor}{black}
    \colorlet{innerblockbodybgcolor}{uofgsandstone}
    \colorlet{innerblockbodyfgcolor}{black}
    % Note colors
    \colorlet{notefgcolor}{black}
    \colorlet{notebgcolor}{uofgrust}
    \colorlet{noteframecolor}{red}
}

\usetheme{Autumn}
\usecolorstyle{UofG}

\tikzposterlatexaffectionproofoff

\useblockstyle[bodyverticalshift=-1cm, roundedcorners=1]{Default}

\renewcommand{\Huge}{\fontsize{77.2}{96}\selectfont}

% Styles for drawings

\tikzset{edge/.style={line width=3pt, color=uofgsandstone}}
\tikzset{ledge/.style={line width=3pt, color=uofgsandstone!40!white}}
\tikzset{hedge/.style={line width=3pt, color=uofgsandstone, dashed}}

\setlength\intextsep{0pt}

\begin{document}
\maketitle

{
    \colorlet{blockbodybgcolor}{uofgcobalt}
    \colorlet{blocktitlebgcolor}{uofgcobalt}
    \block[bodyverticalshift=0cm, bodyinnersep=3mm]{}{
        \centering\begin{minipage}{0.94\textwidth}
            We show how to generate \textbf{really hard} random instances for \textbf{subgraph
            isomorphism} problems. For the non-induced variant, we predict and observe a phase
            transition between satisfiable and unsatisfiable instances, with a corresponding
            complexity peak seen in \textbf{three different solvers}. For the induced variant, much
            richer behaviour is observed, and \textbf{constrainedness} gives a better measure of
            difficulty than does proximity to a phase transition. We also discuss \textbf{variable
            and value ordering} heuristics, and their relationship to the \textbf{expected number of
            solutions}.
        \end{minipage}
    }
}

\begin{columns}
\column{0.5}

\block{Subgraph Isomorphism (Finding Patterns in Graphs)}{
\begin{wrapfigure}[11]{r}{0.46\linewidth}
    \begin{center}\begin{tikzpicture}[scale=1.1]%{{{
        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (Na) at (1,  0)
        {\vphantom{0}};
        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (Nb) at (1, -2)
        {\vphantom{0}};
        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (Nc) at (0, -4)
        {\vphantom{0}};
        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (Nd) at (2, -4)
        {\vphantom{0}};

        \draw [edge] (Na) -- (Nb);
        \draw [edge] (Nb) -- (Nc);
        \draw [edge] (Nc) -- (Nd);
        \draw [edge] (Nb) -- (Nd);

        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (N1) at (7.5,  0) {\vphantom{0}};
        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (N2) at (9.5,  0) {\vphantom{0}};
        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (N3) at (7.5, -2) {\vphantom{0}};
        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (N4) at (9.5, -2) {\vphantom{0}};
        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (N5) at (7.5, -4) {\vphantom{0}};
        \node[draw, circle, fill=uofgrose, inner sep=5pt, font=\bfseries] (N6) at (9.5, -4) {\vphantom{0}};
        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (N7) at (5.5,  0) {\vphantom{0}};
        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (N8) at (5.5, -2) {\vphantom{0}};
        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (N9) at (5.5, -4) {\vphantom{0}};

        \draw [ledge] (N1) -- (N2);
        \draw [edge] (N1) -- (N3);
        \draw [ledge] (N1) -- (N4);
        \draw [ledge] (N2) -- (N4);
        \draw [edge] (N3) -- (N5);
        \draw [edge] (N3) -- (N6);
        \draw [ledge] (N4) -- (N6);
        \draw [edge] (N5) -- (N6);
        \draw [ledge] (N2) to [in=45, out=315] (N6);
        \draw [ledge] (N1) -- (N7);
        \draw [ledge] (N5) -- (N9);
        \draw [ledge] (N7) -- (N8);
        \draw [ledge] (N8) -- (N9);
    \end{tikzpicture}\end{center}

    \vspace*{1.5cm}

    \begin{center}\begin{tikzpicture}[scale=1.1]%{{{
        \node[draw, circle, fill=uofgcobalt, inner sep=5pt, font=\bfseries] (Na) at (1,  0)
        {\vphantom{0}};
        \node[draw, circle, fill=uofgcobalt, inner sep=5pt, font=\bfseries] (Nb) at (1, -2)
        {\vphantom{0}};
        \node[draw, circle, fill=uofgcobalt, inner sep=5pt, font=\bfseries] (Nc) at (0, -4)
        {\vphantom{0}};
        \node[draw, circle, fill=uofgcobalt, inner sep=5pt, font=\bfseries] (Nd) at (2, -4)
        {\vphantom{0}};

        \draw [edge] (Na) -- (Nb);
        \draw [edge] (Nb) -- (Nc);
        \draw [edge] (Nc) -- (Nd);
        \draw [edge] (Nb) -- (Nd);

        \node[draw, circle, fill=uofgcobalt, inner sep=5pt, font=\bfseries] (N1) at (7.5,  0) {\vphantom{0}};
        \node[draw, circle, fill=uofgcobalt, inner sep=5pt, font=\bfseries] (N2) at (9.5,  0) {\vphantom{0}};
        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (N3) at (7.5, -2) {\vphantom{0}};
        \node[draw, circle, fill=uofgcobalt, inner sep=5pt, font=\bfseries] (N4) at (9.5, -2) {\vphantom{0}};
        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (N5) at (7.5, -4) {\vphantom{0}};
        \node[draw, circle, fill=uofgcobalt, inner sep=5pt, font=\bfseries] (N6) at (9.5, -4) {\vphantom{0}};
        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (N7) at (5.5,  0) {\vphantom{0}};
        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (N8) at (5.5, -2) {\vphantom{0}};
        \node[draw, circle, fill=white, inner sep=5pt, font=\bfseries] (N9) at (5.5, -4) {\vphantom{0}};

        \draw [edge] (N1) -- (N2);
        \draw [ledge] (N1) -- (N3);
        \draw [edge] (N1) -- (N4);
        \draw [edge] (N2) -- (N4);
        \draw [ledge] (N3) -- (N5);
        \draw [ledge] (N3) -- (N6);
        \draw [edge] (N4) -- (N6);
        \draw [ledge] (N5) -- (N6);
        \draw [hedge] (N2) to [in=45, out=315] (N6);
        \draw [ledge] (N1) -- (N7);
        \draw [ledge] (N5) -- (N9);
        \draw [ledge] (N7) -- (N8);
        \draw [ledge] (N8) -- (N9);
    \end{tikzpicture}\end{center}
\end{wrapfigure}

The \textbf{non-induced subgraph isomorphism problem} is to find an injective mapping from a given
pattern graph to a given target graph which preserves adjacency---in essence, we are ``finding a
copy of'' the pattern inside the target.

\medskip

The \textbf{induced} variant of the problem additionally requires that the mapping preserve
non-adjacency, so there are no ``extra edges'' in the copy of the pattern that we find. The top
example is induced, whereas the bottom example is not, due to the dashed edge.

\medskip

Despite these problems being \NP-complete, modern practical subgraph isomorphism algorithms can
handle problem instances with many hundreds of vertices in the pattern graph, and up to \textbf{ten thousand
vertices} in the target graph, leading to successful application in areas such as \textbf{computer
vision}, \textbf{biochemistry}, and \textbf{pattern recognition}. However, these algorithms cannot
handle arbitrary instances of this size---here we generate instances with thirty and one hundred and
fifty vertices respectively which cannot be solved within one hour.
}

\block{The Expected Number of Solutions, and Heuristics}{
    If we randomly generate a pattern graph with $p$ vertices and density $d_p$, and a target graph
    with $t$ vertices and density $d_t$, the expected number of solutions to the non-induced
    isomorphism problem is \[
        \mathsf{\langle Sol \rangle = t \cdot (t - 1) \cdot \ldots \cdot (t - p + 1)\cdot {d_t}^{d_p \cdot
        \binom{p}{2}} } \textnormal{.} \]
}

\column{0.5}

\block{Non-Induced Phase Transitions}{
    If we generate random patterns and targets independently, fixing a pattern order of 20 vertices,
    a target order of 150 vertices, a target density of 0.4, and varying the pattern density, the
    familiar phase transition and (solver-independent) complexity peak occurs.

    \vspace*{1em}

    \begin{center}
        \small
        \input{gen-graph-phase-transition}
    \end{center}

    We can also look at what happens if we vary both the pattern and the target density, for
    different orders of pattern. The top row shows satisfiability---the black lines show where the
    expected number of solutions is one. The bottom row shows the search cost for one of the three
    algorithms we consider in the paper. As expected, the ``really hard'' instances are near the
    phase transition.

    \vspace*{1em}

    \begin{center}
        \small
        \input{gen-graph-non-induced}
    \end{center}
}

\end{columns}

\block{Induced Phase Transitions}{
\begin{wrapfigure}[21]{r}{0.62\linewidth}\begin{center}
    \small\input{gen-graph-induced}
\end{center}\end{wrapfigure}

    For induced isomorphisms, the expected number of solutions is \[
        \mathsf{\langle Sol \rangle = t \cdot (t - 1) \cdot \ldots \cdot (t - p + 1)\cdot
                    {d_t}^{d_p \cdot \binom{p}{2}} \cdot {(1 - d_{t})}^{(1 - d_{p}) \cdot \binom{p}{2}} } \textnormal{.} \]

    \vspace*{52ex}
}

\begin{columns}
\column{0.5}

\block{See the Paper For\ldots}{
    \begin{itemize}
        \item ~ Results from two other subgraph isomorphism algorithms.
        \item ~ Behaviour under reduction to clique, pseudo-boolean, and boolean satisfiability.
        \item ~ Should we alter heuristics based upon the expected number of solutions?
    \end{itemize}
}

\column{0.5}

\block{Future Work}{
    \begin{itemize}
        \item ~ Variance for more accurate predictions (if you can help, please get in touch).
        \item ~ Other random models (degree-bounded, scale-free, \ldots).
        \item ~ Labelled graphs.
    \end{itemize}
}

\end{columns}

{
    \colorlet{blockbodybgcolor}{uofgsandstone!80!white}
    \colorlet{blocktitlebgcolor}{uofgsandstone!80!white}
    \block[bodyverticalshift=-0.5cm]{}{
        This work was supported by the Engineering and Physical Sciences Research Council [grant
        number EP/K503058/1]. \hfill \texttt{c.mccreesh.1@research.gla.ac.uk}
    }
}

\end{document}