### 2016-06-29 talk/talk.tex
Start IJCAI talk

% vim: set spell spelllang=en tw=100 et sw=4 sts=4 foldmethod=marker foldmarker={{{,}}} :

\documentclass{beamer}

\usepackage{tikz}
\usepackage{xcolor}
\usepackage{complexity}
\usepackage{hyperref}
\usepackage{microtype}
\usepackage{amsmath}                   % \operatorname
\usepackage{amsfonts}                  % \mathcal
\usepackage{amssymb}                   % \nexists
\usepackage{gnuplot-lua-tikz}          % graphs
\usepackage[vlined]{algorithm2e} % algorithms
\usepackage{centernot}

\usetikzlibrary{shapes, arrows, shadows, calc, positioning, fit}
\usetikzlibrary{decorations.pathreplacing, decorations.pathmorphing, shapes.misc}
\usetikzlibrary{tikzmark}

\definecolor{uofguniversityblue}{rgb}{0, 0.219608, 0.396078}

\definecolor{uofgheather}{rgb}{0.356863, 0.32549, 0.490196}
\definecolor{uofgaquamarine}{rgb}{0.603922, 0.72549, 0.678431}
\definecolor{uofgslate}{rgb}{0.309804, 0.34902, 0.380392}
\definecolor{uofgrose}{rgb}{0.823529, 0.470588, 0.709804}
\definecolor{uofgmocha}{rgb}{0.709804, 0.564706, 0.47451}
\definecolor{uofgsandstone}{rgb}{0.321569, 0.278431, 0.231373}
\definecolor{uofgforest}{rgb}{0, 0.2, 0.129412}
\definecolor{uofglawn}{rgb}{0.517647, 0.741176, 0}
\definecolor{uofgcobalt}{rgb}{0, 0.615686, 0.92549}
\definecolor{uofgturquoise}{rgb}{0, 0.709804, 0.819608}
\definecolor{uofgsunshine}{rgb}{1.0, 0.862745, 0.211765}
\definecolor{uofgpumpkin}{rgb}{1.0, 0.72549, 0.282353}
\definecolor{uofgthistle}{rgb}{0.584314, 0.070588, 0.447059}
\definecolor{uofgrust}{rgb}{0.603922, 0.227451, 0.023529}
\definecolor{uofgburgundy}{rgb}{0.490196, 0.133333, 0.223529}
\definecolor{uofgpillarbox}{rgb}{0.701961, 0.047059, 0}
\definecolor{uofglavendar}{rgb}{0.356863, 0.301961, 0.580392}

\tikzset{vertex/.style={draw, circle, inner sep=0pt, minimum size=0.5cm, font=\small\bfseries}}
\tikzset{notvertex/.style={vertex, color=white, text=black}}
\tikzset{plainvertex/.style={vertex}}
\tikzset{vertexc1/.style={vertex, fill=uofgburgundy, text=white}}
\tikzset{vertexc2/.style={vertex, fill=uofgsandstone, text=white}}
\tikzset{vertexc3/.style={vertex, fill=uofgforest, text=white}}
\tikzset{vertexc4/.style={vertex, fill=uofgheather, text=white}}
\tikzset{edge/.style={color=black!50!white}}
\tikzset{bedge/.style={ultra thick}}

% {{{ theme things
\useoutertheme[footline=authortitle]{miniframes}
\useinnertheme{rectangles}

\setbeamerfont{block title}{size={}}
\setbeamerfont{title}{size=\large,series=\bfseries}
\setbeamerfont{section title}{size=\large,series=\mdseries}
\setbeamerfont{author}{size=\normalsize,series=\mdseries}
\setbeamercolor*{structure}{fg=uofguniversityblue}
\setbeamercolor*{palette primary}{use=structure,fg=black,bg=white}
\setbeamercolor*{palette secondary}{use=structure,fg=white,bg=uofgcobalt}
\setbeamercolor*{palette tertiary}{use=structure,fg=white,bg=uofguniversityblue}
\setbeamercolor*{palette quaternary}{fg=white,bg=black}

\setbeamercolor*{titlelike}{parent=palette primary}

\beamertemplatenavigationsymbolsempty

\setbeamertemplate{title page}
{
    \begin{tikzpicture}[remember picture, overlay]
        \node at (current page.north west) {
            \begin{tikzpicture}[remember picture, overlay]
                \fill [fill=uofguniversityblue, anchor=north west] (0, 0) rectangle (\paperwidth, -2.6cm);
            \end{tikzpicture}
        };

        \node (logo) [anchor=north east, shift={(-0.6cm,-0.6cm)}] at (current page.north east) {
            \includegraphics*[keepaspectratio=true,scale=0.7]{UoG_keyline.pdf}
        };

        \node [anchor=west, xshift=0.2cm] at (current page.west |- logo.west) {
            \begin{minipage}{0.65\paperwidth}\raggedright
                {\usebeamerfont{title}\usebeamercolor[white]{}\inserttitle}\\[0.1cm]
                {\usebeamerfont{author}\usebeamercolor[white]{}\insertauthor}
            \end{minipage}
        };
    \end{tikzpicture}
}

\setbeamertemplate{section page}
{
    \begin{centering}
        \begin{beamercolorbox}[sep=12pt,center]{part title}
            \usebeamerfont{section title}\insertsection\par
        \end{beamercolorbox}
    \end{centering}
}

\newcommand{\frameofframes}{/}
\newcommand{\setframeofframes}[1]{\renewcommand{\frameofframes}{#1}}

\makeatletter
\setbeamertemplate{footline}
{%
    \begin{beamercolorbox}[colsep=1.5pt]{upper separation line foot}
    \end{beamercolorbox}
    \begin{beamercolorbox}[ht=2.5ex,dp=1.125ex,%
        leftskip=.3cm,rightskip=.3cm plus1fil]{author in head/foot}%
        \leavevmode{\usebeamerfont{author in head/foot}\insertshortauthor}%
        \hfill%
        {\usebeamerfont{institute in head/foot}\usebeamercolor[fg]{institute in head/foot}\insertshortinstitute}%
    \end{beamercolorbox}%
    \begin{beamercolorbox}[ht=2.5ex,dp=1.125ex,%
        leftskip=.3cm,rightskip=.3cm plus1fil]{title in head/foot}%
        {\usebeamerfont{title in head/foot}\insertshorttitle}%
        \hfill%
        {\usebeamerfont{frame number}\usebeamercolor[fg]{frame number}\insertframenumber~\frameofframes~\inserttotalframenumber}
    \end{beamercolorbox}%
    \begin{beamercolorbox}[colsep=1.5pt]{lower separation line foot}
    \end{beamercolorbox}
}

\makeatletter
\newenvironment{nearlyplainframe}[2][]{
    \def\beamer@entrycode{\vspace*{-\headheight}}
    \setbeamertemplate{headline}
    {%
        \begin{beamercolorbox}[colsep=1.5pt]{upper separation line head}
        \end{beamercolorbox}
        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
            leftskip=.3cm,rightskip=.3cm plus1fil]{title in head/foot}%
        \end{beamercolorbox}%
        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
            leftskip=.3cm,rightskip=.3cm plus1fil]{author in head/foot}%
        \end{beamercolorbox}%
        \begin{beamercolorbox}[colsep=1.5pt]{lower separation line head}
        \end{beamercolorbox}
        \vspace*{\headheight}
    }

    \setbeamertemplate{footline}
    {%
        \begin{beamercolorbox}[colsep=1.5pt]{upper separation line foot}
        \end{beamercolorbox}
        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
            leftskip=.3cm,rightskip=.3cm plus1fil]{author in head/foot}%
        \end{beamercolorbox}%
        \begin{beamercolorbox}[ht=0.5ex,dp=0.125ex,%
            leftskip=.3cm,rightskip=.3cm plus1fil]{title in head/foot}%
        \end{beamercolorbox}%
        \begin{beamercolorbox}[colsep=1.5pt]{lower separation line foot}
        \end{beamercolorbox}
    }

    \begin{frame}[#1]{#2}
    }{
    \end{frame}
}
\makeatother

% }}}

\title[Heuristics and Really Hard Instances for Subgraph Isomorphism Problems]{Heuristics and Really
Hard Instances \\ for Subgraph Isomorphism Problems}
\author[Ciaran McCreesh, Patrick Prosser and James Trimble]{\textbf{Ciaran McCreesh}, Patrick
Prosser and \\ James Trimble}

\begin{document}

{
    \usebackgroundtemplate{
        \tikz[overlay, remember picture]
        \node[at=(current page.south), anchor=south, inner sep=0pt]{\includegraphics*[keepaspectratio=true, width=\paperwidth]{background.jpg}};
    }
    \begin{frame}[plain,noframenumbering]
        \titlepage
    \end{frame}
}

\begin{frame}{Non-Induced Subgraph Isomorphism}

    \begin{center}\begin{tikzpicture}
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Na) at (1,  0) {1};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Nb) at (1, -1.5) {2};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Nc) at (0, -3) {3};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Nd) at (2, -3) {4};

        \node <2> [draw, circle, fill=uofgcobalt, inner sep=4pt, font=\bfseries] (Na) at (1,  0) {1};
        \node <2> [draw, circle, fill=uofgpumpkin, inner sep=4pt, font=\bfseries] (Nb) at (1, -1.5) {2};
        \node <2> [draw, circle, fill=uofglawn, inner sep=4pt, font=\bfseries] (Nc) at (0, -3) {3};
        \node <2> [draw, circle, fill=uofgthistle, inner sep=4pt, font=\bfseries] (Nd) at (2, -3) {4};

        \draw <1> [thick, color=uofgsandstone!50] (Na) -- (Nb);
        \draw <1> [thick, color=uofgsandstone!50] (Nb) -- (Nc);
        \draw <1> [thick, color=uofgsandstone!50] (Nc) -- (Nd);
        \draw <1> [thick, color=uofgsandstone!50] (Nb) -- (Nd);

        \draw <2> [ultra thick] (Na) -- (Nb);
        \draw <2> [ultra thick] (Nb) -- (Nc);
        \draw <2> [ultra thick] (Nc) -- (Nd);
        \draw <2> [ultra thick] (Nb) -- (Nd);

        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N1) at (5.5,  0) {1};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N2) at (7.5,  0) {2};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N3) at (5.5, -1.5) {3};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N4) at (7.5, -1.5) {4};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N5) at (5.5, -3) {5};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N6) at (7.5, -3) {6};

        \node <2> [draw, circle, fill=uofgcobalt, inner sep=4pt, font=\bfseries] (N1) at (5.5,  0) {1};
        \node <2> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N2) at (7.5,  0) {2};
        \node <2> [draw, circle, fill=uofgpumpkin, inner sep=4pt, font=\bfseries] (N3) at (5.5, -1.5) {3};
        \node <2> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N4) at (7.5, -1.5) {4};
        \node <2> [draw, circle, fill=uofglawn, inner sep=4pt, font=\bfseries] (N5) at (5.5, -3) {5};
        \node <2> [draw, circle, fill=uofgthistle, inner sep=4pt, font=\bfseries] (N6) at (7.5, -3) {6};

        \draw <1> [thick, color=uofgsandstone!50] (N1) -- (N2);
        \draw <1> [thick, color=uofgsandstone!50] (N1) -- (N3);
        \draw <1> [thick, color=uofgsandstone!50] (N1) -- (N4);
        \draw <1> [thick, color=uofgsandstone!50] (N2) -- (N4);
        \draw <1> [thick, color=uofgsandstone!50] (N3) -- (N5);
        \draw <1> [thick, color=uofgsandstone!50] (N3) -- (N6);
        \draw <1> [thick, color=uofgsandstone!50] (N4) -- (N6);
        \draw <1> [thick, color=uofgsandstone!50] (N5) -- (N6);
        \draw <1> [thick, color=uofgsandstone!50] (N2) to [in=45, out=315] (N6);
        \draw <1> [thick, color=uofgsandstone!50] (N1) to [in=135, out=225] (N5);

        \draw <2> [thick, color=uofgsandstone!50] (N1) -- (N2);
        \draw <2> [ultra thick] (N1) -- (N3);
        \draw <2> [thick, color=uofgsandstone!50] (N1) -- (N4);
        \draw <2> [thick, color=uofgsandstone!50] (N2) -- (N4);
        \draw <2> [ultra thick] (N3) -- (N5);
        \draw <2> [ultra thick] (N3) -- (N6);
        \draw <2> [thick, color=uofgsandstone!50] (N4) -- (N6);
        \draw <2> [ultra thick] (N5) -- (N6);
        \draw <2> [thick, color=uofgsandstone!50] (N2) to [in=45, out=315] (N6);
        \draw <2> [thick, color=uofgsandstone!50] (N1) to [in=135, out=225] (N5);
    \end{tikzpicture}\end{center}

\end{frame}

\begin{frame}{Induced Subgraph Isomorphism}

    \begin{center}\begin{tikzpicture}
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Na) at (1,  0) {1};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Nb) at (1, -1.5) {2};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Nc) at (0, -3) {3};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (Nd) at (2, -3) {4};

        \node <2> [draw, circle, fill=uofgcobalt, inner sep=4pt, font=\bfseries] (Na) at (1,  0) {1};
        \node <2> [draw, circle, fill=uofgpumpkin, inner sep=4pt, font=\bfseries] (Nb) at (1, -1.5) {2};
        \node <2> [draw, circle, fill=uofglawn, inner sep=4pt, font=\bfseries] (Nc) at (0, -3) {3};
        \node <2> [draw, circle, fill=uofgthistle, inner sep=4pt, font=\bfseries] (Nd) at (2, -3) {4};

        \draw <1> [thick, color=uofgsandstone!50] (Na) -- (Nb);
        \draw <1> [thick, color=uofgsandstone!50] (Nb) -- (Nc);
        \draw <1> [thick, color=uofgsandstone!50] (Nc) -- (Nd);
        \draw <1> [thick, color=uofgsandstone!50] (Nb) -- (Nd);

        \draw <2> [ultra thick] (Na) -- (Nb);
        \draw <2> [ultra thick] (Nb) -- (Nc);
        \draw <2> [ultra thick] (Nc) -- (Nd);
        \draw <2> [ultra thick] (Nb) -- (Nd);

        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N1) at (5.5,  0) {1};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N2) at (7.5,  0) {2};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N3) at (5.5, -1.5) {3};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N4) at (7.5, -1.5) {4};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N5) at (5.5, -3) {5};
        \node <1> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N6) at (7.5, -3) {6};

        \node <2> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N1) at (5.5,  0) {1};
        \node <2> [draw, circle, fill=uofgcobalt, inner sep=4pt, font=\bfseries] (N2) at (7.5,  0) {2};
        \node <2> [draw, circle, fill=uofgthistle, inner sep=4pt, font=\bfseries] (N3) at (5.5, -1.5) {3};
        \node <2> [draw, circle, fill=white, inner sep=4pt, font=\bfseries] (N4) at (7.5, -1.5) {4};
        \node <2> [draw, circle, fill=uofglawn, inner sep=4pt, font=\bfseries] (N5) at (5.5, -3) {5};
        \node <2> [draw, circle, fill=uofgpumpkin, inner sep=4pt, font=\bfseries] (N6) at (7.5, -3) {6};

        \draw <1> [thick, color=uofgsandstone!50] (N1) -- (N2);
        \draw <1> [thick, color=uofgsandstone!50] (N1) -- (N3);
        \draw <1> [thick, color=uofgsandstone!50] (N1) -- (N4);
        \draw <1> [thick, color=uofgsandstone!50] (N2) -- (N4);
        \draw <1> [thick, color=uofgsandstone!50] (N3) -- (N5);
        \draw <1> [thick, color=uofgsandstone!50] (N3) -- (N6);
        \draw <1> [thick, color=uofgsandstone!50] (N4) -- (N6);
        \draw <1> [thick, color=uofgsandstone!50] (N5) -- (N6);
        \draw <1> [thick, color=uofgsandstone!50] (N2) to [in=45, out=315] (N6);
        \draw <1> [thick, color=uofgsandstone!50] (N1) to [in=135, out=225] (N5);

        \draw <2> [thick, color=uofgsandstone!50] (N1) -- (N2);
        \draw <2> [thick, color=uofgsandstone!50] (N1) -- (N3);
        \draw <2> [thick, color=uofgsandstone!50] (N1) -- (N4);
        \draw <2> [thick, color=uofgsandstone!50] (N2) -- (N4);
        \draw <2> [ultra thick] (N3) -- (N5);
        \draw <2> [ultra thick] (N3) -- (N6);
        \draw <2> [thick, color=uofgsandstone!50] (N4) -- (N6);
        \draw <2> [ultra thick] (N5) -- (N6);
        \draw <2> [ultra thick] (N2) to [in=45, out=315] (N6);
        \draw <2> [thick, color=uofgsandstone!50] (N1) to [in=135, out=225] (N5);
    \end{tikzpicture}\end{center}
\end{frame}

\begin{frame}{Benchmarking}

    \begin{itemize}
        \item Based upon chemical and computer vision datasets, we can handle patterns with 1,000
            vertices and targets with 10,000 vertices.
        \item Do these results reflect the worst case, or are they too optimistic?
        \item Can we create ``hard'' benchmark instances?
    \end{itemize}

\end{frame}

\begin{frame}{Randomly Selected Subgraphs}

    \begin{itemize}
        \item Start with a random target graph.
        \item Pick vertices at random to make a pattern.
        \item Shuffle the numbering.
        \item <2-> These instances will always be satisfiable!
    \end{itemize}

\end{frame}

\begin{frame}{Independently Random Subgraphs}

    \begin{itemize}
        \item Make a random target graph.
        \item Independently, make a random pattern graph.
        \item <2-> Will these instances ever be satisfiable?
    \end{itemize}

\end{frame}

\begin{nearlyplainframe}[t]{A Phase Transition}

    \input{gen-graph-phase}

\end{nearlyplainframe}

\begin{frame}{This Looks Familiar\ldots}
    \centering
    \includegraphics*[keepaspectratio=true,scale=0.25]{sat.jpg}

    \par\flushleft
    Understanding the Empirical Hardness of NP-Complete Problems. Kevin Leyton-Brown, Holger H.
    Hoos, Frank Hutter, Lin Xu.  Communications of the ACM, Vol. 57 No. 5, Pages 98-107
\end{frame}

\begin{nearlyplainframe}[t]{In Two Dimensions?}

\end{nearlyplainframe}

\begin{frame}{See The Paper For\ldots}

    \begin{itemize}
        \item Is this behaviour solver-independent?
        \item Estimating the phase transition location
        \item Using this to rediscover variable and value ordering heuristics
    \end{itemize}

\end{frame}

\begin{nearlyplainframe}[t]{What About Induced?}

\end{nearlyplainframe}

\begin{nearlyplainframe}[t]{What Changes Between 10 and 20?}

\end{nearlyplainframe}

\begin{nearlyplainframe}[t]{Is The Central Region Really Hard?}

\end{nearlyplainframe}

\begin{nearlyplainframe}[t]{Constrainedness}

\end{nearlyplainframe}

\begin{frame}{See The Paper For\ldots}

    \begin{itemize}
        \item Is this behaviour solver-independent?
            \begin{itemize}
                \item Also SAT, PB, MIP encodings, reduction to clique.
            \end{itemize}
        \item Estimating the phase transition location
        \item Using this to invent new variable and value ordering heuristics
            \begin{itemize}
                \item But something unexpected happens this time!
            \end{itemize}
    \end{itemize}

\end{frame}

\begin{frame}{Future Work}

    \begin{itemize}
        \item Other randomness models (bounded degree, regular, scale-free).
        \item Better estimates of the phase transition location for very sparse or very dense
            patterns.
            \begin{itemize}
                \item This needs a horrible variance calculation. Please get in touch if you like
                    doing this sort of thing.
            \end{itemize}
        \item Dynamic heuristics?
    \end{itemize}

\end{frame}

\begin{frame}[plain,noframenumbering]
    \begin{center}
        \vspace*{4em}
        \url{http://www.dcs.gla.ac.uk/~ciaran} \\
        \vspace{1em}
        \href{mailto:c.mccreesh.1@research.gla.ac.uk}{\nolinkurl{c.mccreesh.1@research.gla.ac.uk}} \\
    \end{center}

    \begin{tikzpicture}[remember picture, overlay]
        \node at (current page.north west) {
            \begin{tikzpicture}[remember picture, overlay]
                \fill [fill=uofguniversityblue, anchor=north west] (0, 0) rectangle (\paperwidth, -2.6cm);
            \end{tikzpicture}
        };

        \node (logo) [anchor=north east, shift={(-0.6cm,-0.6cm)}] at (current page.north east) {
            \includegraphics*[keepaspectratio=true,scale=0.7]{UoG_keyline.pdf}
        };
    \end{tikzpicture}
\end{frame}

\end{document}